---
title: "Analysis for Figure 3"
author: "MD"
date: "`r Sys.Date()`"
output: 
 html_document:
   toc: true
   toc_float: true
   editor_options:
   chunck_output_type: console
---

Inputs: <br>

-   Sequencing_data_formatted_for_analysis(01.28.2026).xlsx
-   EJM-CLS-22F_pct98.csv
-   EJM-CLS-22B_pct98.csv
-   sample_data_field_switch_samples_bac.csv
 

Outputs: <br>

-   figure 3
-   ancom_bac_farm.rds
-   ancom_fun_farm.rds

# Load libraries

```{r, results = 'hide', message=FALSE}
knitr::opts_chunk$set(out.width = "100%")
library(ANCOMBC)
library(tidyverse)
library(readxl)
library(phyloseq)
library(mia)
library(cowplot)
```

# Generate phyloseq and tree summarized object

```{r}
FUN <- read.csv("../Data/EJM-CLS-22F_pct98.csv")
BAC <- read.csv("../Data/EJM-CLS-22B_pct98.csv")
BAC <- BAC[!is.na(BAC$Species),]
FUN <- FUN[FUN$Kingdom =="k__Fungi",]
FUN <- FUN[!is.na(FUN$Species),]
SAMPLE_DATA <- as.data.frame(read_xlsx("../Data/Sequencing_data_formatted_for_analysis(01.28.2026).xlsx", sheet = "Soil_sample_info"))
rownames(FUN) <- FUN$X
FUN <- FUN[,-1]
rownames(BAC) <- BAC$X
BAC <- BAC[,-1]
BAC$Species <- paste(BAC$Genus,BAC$Species)
FUN$Species <- paste(FUN$Genus,FUN$Species)
FUN$Species <- gsub("s__", " ", FUN$Species)
FUN$Species <- gsub("g__", "", FUN$Species)

SAMPLE_DATA$EJM_SOIL_NUMBER <- as.factor(SAMPLE_DATA$EJM_SOIL_NUMBER)

FUN_OTU<- FUN[,8:length(colnames(FUN))]
FUN_TAX <- FUN[,1:7]
BAC_OTU<- BAC[,8:length(colnames(BAC))]
BAC_TAX <- BAC[,1:7]

BAC_OTU <- BAC_OTU[rowSums(BAC_OTU)>1,]
FUN_OTU <- FUN_OTU[rowSums(FUN_OTU)>1,]

BAC_OTU <- BAC_OTU[,colSums(BAC_OTU)>1000]
FUN_OTU <- FUN_OTU[,colSums(FUN_OTU)>1000]

colSums(BAC_OTU)

rownames(SAMPLE_DATA)<- SAMPLE_DATA$SEQUENCED_INV_NUMBER
```

## Bacteria

```{r, message = F}
SAMPLE_DATA_FIELD <- SAMPLE_DATA[SAMPLE_DATA$SOIL_SOURCE=="FIELD",]
OTU <- BAC_OTU[,colnames(BAC_OTU) %in% rownames(SAMPLE_DATA_FIELD)]
OTU <- OTU[rowSums(OTU)>1,]
OTU <- as.matrix(OTU)
TAX = BAC_TAX[rownames(OTU),]
samples = SAMPLE_DATA_FIELD[colnames(OTU),]


samples <- read.csv("../Data/output/sample_data_field_switch_samples_bac.csv")
samples <- samples %>%
mutate(Farm = as_factor(Farm)) %>%
mutate(Farm = fct_relevel(Farm, "CROSS", "TT", "REM", "NMF"))

sample_names(samples) <- samples$SEQUENCED_INV_NUMBER
rownames(samples) <- samples$SEQUENCED_INV_NUMBER


em_tse <- TreeSummarizedExperiment(assays = list(Count = OTU),
                                   rowData = TAX,
                                   colData = samples)
assayNames(em_tse) <- "counts" #Needed for prevalence filtering
P.bac <- convertToPhyloseq(em_tse,assay.type = "counts")
```

## Fungi

```{r, message = F}
SAMPLE_DATA_FIELD <- SAMPLE_DATA[SAMPLE_DATA$SOIL_SOURCE=="FIELD",]
OTU <- FUN_OTU[,colnames(FUN_OTU) %in% rownames(SAMPLE_DATA_FIELD)]
OTU <- OTU[rowSums(OTU)>1,]
OTU <- as.matrix(OTU)
TAX = FUN_TAX[rownames(OTU),]
samples = SAMPLE_DATA_FIELD[colnames(OTU),]


samples <- read.csv("../Data/output/sample_data_field_switch_samples_fun.csv")
samples <- samples %>%
mutate(Farm = as_factor(Farm)) %>%
mutate(Farm = fct_relevel(Farm, "CROSS", "TT", "REM", "NMF"))


sample_names(samples) <- samples$SEQUENCED_INV_NUMBER
rownames(samples) <- samples$SEQUENCED_INV_NUMBER
em_tse <- TreeSummarizedExperiment(assays = list(Count = OTU),
                                   rowData = TAX,
                                   colData = samples)
assayNames(em_tse) <- "counts" #Needed for prevalence filtering
P.fun <- convertToPhyloseq(em_tse,assay.type = "counts")
```

## Subset to farms of interest

```{r}
P.sub.bac <- subset_samples(P.bac, Farm == "TT" | Farm == "CROSS")
bac_farm_tse = mia::convertFromPhyloseq(P.sub.bac)
P.sub.fun <- subset_samples(P.fun, Farm == "TT" | Farm == "CROSS")
fun_farm_tse = mia::convertFromPhyloseq(P.sub.fun)
```

# Runs ancombc

## Bacteria

```{r}
set.seed(123)
Sys.time()
ancom_bac_SPE = ancombc2(data = bac_farm_tse, assay_name = "counts",tax_level = "Species",
                                 fix_formula = "Farm",
                                 p_adj_method = "holm", pseudo_sens = TRUE,
                                 prv_cut = 0.10, lib_cut = 0, s0_perc = 0.05,
                                 group = NULL, struc_zero = FALSE, neg_lb = FALSE,
                                 alpha = 0.05, n_cl = 4, verbose = TRUE,
                                 global = FALSE, pairwise = FALSE, dunnet = FALSE, trend = FALSE,
                                 iter_control = list(tol = 1e-2, max_iter = 20, verbose = TRUE),
                                 em_control = list(tol = 1e-5, max_iter = 100),
                                 lme_control = lme4::lmerControl(),
                                 mdfdr_control = list(fwer_ctrl_method = "holm", B = 100),
                                 trend_control = NULL)
Sys.time()
saveRDS(ancom_bac_SPE, file = '../Data/output/ancom_bac_farm.rds', ascii = FALSE, version = NULL, compress = TRUE, refhook = NULL)
```

## Fungi

```{r}
set.seed(123)
Sys.time()
ancom_fun_SPE = ancombc2(data = fun_farm_tse, assay_name = "counts",tax_level = "Species",
                                 fix_formula = "Farm",
                                 p_adj_method = "holm", pseudo_sens = TRUE,
                                 prv_cut = 0.10, lib_cut = 0, s0_perc = 0.05,
                                 group = NULL, struc_zero = FALSE, neg_lb = FALSE,
                                 alpha = 0.05, n_cl = 4, verbose = TRUE,
                                 global = FALSE, pairwise = FALSE, dunnet = FALSE, trend = FALSE,
                                 iter_control = list(tol = 1e-2, max_iter = 20, verbose = TRUE),
                                 em_control = list(tol = 1e-5, max_iter = 100),
                                 lme_control = lme4::lmerControl(),
                                 mdfdr_control = list(fwer_ctrl_method = "holm", B = 100),
                                 trend_control = NULL)
Sys.time()
saveRDS(ancom_fun_SPE, file = '../Data/output/ancom_fun_farm.rds', ascii = FALSE, version = NULL, compress = TRUE, refhook = NULL)
```

## Read in RDS

```{r}
ancom_bac_SPE <- readRDS(file = '../Data/output/ancom_bac_farm.rds', refhook = NULL)
ancom_fun_SPE <- readRDS(file = '../Data/output/ancom_fun_farm.rds', refhook = NULL)

```

# Generates figure

## Bacteria

### Creates df for figure

```{r}
res_prim = ancom_bac_SPE$res

df = res_prim %>%
  dplyr::filter(p_FarmTT <1) %>%
  dplyr::arrange((lfc_FarmTT)) %>%
  dplyr::filter(passed_ss_FarmTT == 1) %>%
  dplyr::mutate(direct = ifelse(lfc_FarmTT > 0, "Turtle Tree", "Crossroads"))
df$taxon = factor(df$taxon, levels = df$taxon)
df$direct = factor(df$direct,  levels = c("Turtle Tree", "Crossroads"))

df$sig <- ifelse(df$p_FarmTT>0.05, "NS",
                                       ifelse(df$p_FarmTT==0.05, "NS",
                                              ifelse(df$p_FarmTT<0.05, "sig",NA)))
df$sig.stage <- paste(df$direct,df$sig)
df$colors <- ifelse(df$sig=="NS", "NS",
                        ifelse(df$sig.stage=='Turtle Tree sig',"Turtle Tree",
                               ifelse(df$sig.stage=='Crossroads sig','Crossroads',NA)))
df<- arrange(df, -log(p_FarmTT))
df$colors <- factor(df$colors, levels=c('Crossroads', 'Turtle Tree', 'NS'))


df$enrich <- ifelse(df$p_FarmTT < 0.05, paste(df$taxon), NA)
```

### Makes figure

```{r, warning = F}
p.bac<-ggplot(data=df)+
  geom_point(alpha=0.85,mapping= aes(
    fill=colors, x=lfc_FarmTT,y=-log(p_FarmTT)),
    size=5,color='black',pch=21) +
  scale_fill_manual(values = c('#B03A2E','#D4AC0D','gray50'))+
  theme_classic()+
  labs(y=expression("-log"["10"] ~ "p" ["adj"]), 
       x=expression("log"["2"]~"Fold Change"))+
  geom_text(aes(label = enrich,x=lfc_FarmTT,y=-log(p_FarmTT)),hjust=-.1, vjust = -.1,fontface="italic")+
  coord_cartesian(xlim=c(-2.5,5))+
  theme(axis.text=element_text(size=10, colour = "black"), 
        axis.line = element_line(colour = "black"),
        legend.position="none",
        legend.title = element_blank());p.bac

```

## Fungi

### Creates df for figure

```{r}
res_prim = ancom_fun_SPE$res

df = res_prim %>%
  dplyr::filter(p_FarmTT <1) %>%
  dplyr::arrange((lfc_FarmTT)) %>%
  dplyr::filter(passed_ss_FarmTT == 1) %>%
  dplyr::mutate(direct = ifelse(lfc_FarmTT > 0, "Turtle Tree", "Crossroads"))
df$taxon = factor(df$taxon, levels = df$taxon)
df$direct = factor(df$direct,  levels = c("Turtle Tree", "Crossroads"))

df$sig <- ifelse(df$p_FarmTT>0.05, "NS",
                                       ifelse(df$p_FarmTT==0.05, "NS",
                                              ifelse(df$p_FarmTT<0.05, "sig",NA)))
df$sig.stage <- paste(df$direct,df$sig)
df$colors <- ifelse(df$sig=="NS", "NS",
                        ifelse(df$sig.stage=='Turtle Tree sig',"Turtle Tree",
                               ifelse(df$sig.stage=='Crossroads sig','Crossroads',NA)))
df<- arrange(df, -log(p_FarmTT))
df$colors <- factor(df$colors, levels=c('Crossroads', 'Turtle Tree', 'NS'))


df$enrich <- ifelse(df$p_FarmTT < 0.05, paste(df$taxon), NA)
```

### Makes figure

```{r, warning = F}
p.fun<-ggplot(data=df)+
  geom_point(alpha=0.85,mapping= aes(
    fill=colors, x=lfc_FarmTT,y=-log(p_FarmTT)),
    size=5,color='black',pch=21) +
  scale_fill_manual(values = c('#B03A2E','#D4AC0D','gray50'))+
  theme_classic()+
  labs(y=expression("-log"["10"] ~ "p" ["adj"]), 
       x=expression("log"["2"]~"Fold Change"))+
  geom_text(aes(label = enrich,x=lfc_FarmTT,y=-log(p_FarmTT)),hjust=-.1, vjust = -.1,fontface="italic")+
  coord_cartesian(xlim=c(-6,6))+
  theme(axis.text=element_text(size=10, colour = "black"), 
        axis.line = element_line(colour = "black"),
        legend.position="bottom",
        legend.title = element_blank());p.fun

```

## Save figure together

```{r, warning = F}
fig3 <- plot_grid(p.bac, p.fun, ncol = 2, nrow = 1,labels = c("A", "B"),label_size = 20, rel_widths = c(1,1.2));fig3
fig3 <- plot_grid(p.bac, p.fun, ncol = 1, nrow = 2,labels = c("A", "B"),label_size = 20, rel_heights = c(1,1));fig3

ggsave(filename = "../Data/output/Farm_Diff_Abund.svg", plot = fig3, width = 10, height = 8, device = "svg")
```
