---
title: "Fig_2_Analysis"
author: "MD and EM"
date: "`r Sys.Date()`"
output: 
 html_document:
   toc: true
   toc_float: true
   editor_options:
   chunck_output_type: console
---

Inputs:

-   seed_farm_locations.csv (MAP)
-   EJM-CLS-22F_pct98.csv (BETA)
-   EJM-CLS-22B_pct98.csv (BETA)
-   sample_data_field_switch_samples_bac.csv (BETA)
-   sample_data_field_switch_samples_fun.csv (BETA)
-   HATCH_compost_bioassay_data.xlsx (APHID)

Outputs:

-   Figure_2.svg

```{r, results = 'hide', message=FALSE}
library(tidyverse)
library(cowplot)
library(readxl)
library(phyloseq)
library(mia)
library(vegan)
library(car)
library(emmeans)
```

# Map

## Read in data

```{r, warning=F}
site_locations<- read.csv("../Data/seed_farm_locations.csv")
site_locations$farm <- as.factor(site_locations$farm)

new_york <- map_data("county")
new_york<- new_york[new_york$region == "new york",]
new_york$Latitude <- new_york$lat
new_york$Longitude <- new_york$long
```

## Generate figure

```{r, warning=F}
Map <- ggplot() + 
  geom_polygon( data=new_york, aes(x=Longitude, y=Latitude, group = group), fill="lightblue" ) +
  theme_classic()+
  geom_point(data = site_locations, size = 5, aes(x=Longitude, y=Latitude, color = farm))+
  scale_color_manual(values=c("#B03A2E","#21618C","#239B56","#D4AC0D")) +
  theme(legend.position = "none")
MAP <- plot_grid(Map, labels = c('A'));MAP
ggsave(filename = "../Data/output/Seed_farm_location_map.svg", plot = Map, width = 4, height = 4, device = "svg")
```

# Ordination

## Read in data

```{r}
FUN <- read.csv("../Data/EJM-CLS-22F_pct98.csv")
BAC <- read.csv("../Data/EJM-CLS-22B_pct98.csv")
BAC <- BAC[!is.na(BAC$Kingdom),]
FUN <- FUN[FUN$Kingdom =="k__Fungi",]
FUN <- FUN[!is.na(FUN$Kingdom),]
SAMPLE_DATA <- as.data.frame(read_xlsx("../Data/Sequencing_data_formatted_for_analysis(01.28.2026).xlsx", sheet = "Soil_sample_info"))
rownames(FUN) <- FUN$X
FUN <- FUN[,-1]
rownames(BAC) <- BAC$X
BAC <- BAC[,-1]
```

## Create phyloseq

Data preprocessing

```{r}
BAC$Kingdom<-sub("^","k__",BAC$Kingdom)
BAC$Kingdom<-sub("k__NA",NA,BAC$Kingdom)
BAC$Phylum<-sub("^","p__",BAC$Phylum)
BAC$Phylum<-sub("p__NA",NA,BAC$Phylum)
BAC$Class<-sub("^","c__",BAC$Class)
BAC$Class<-sub("c__NA",NA,BAC$Class)
BAC$Order<-sub("^","o__",BAC$Order)
BAC$Order<-sub("o__NA",NA,BAC$Order)
BAC$Family<-sub("^","f__",BAC$Family)
BAC$Family<-sub("f__NA",NA,BAC$Family)
BAC$Genus<-sub("^","g__",BAC$Genus)
BAC$Genus<-sub("g__NA",NA,BAC$Genus)
BAC$Species<-sub("^","s__",BAC$Species)
BAC$Species<-sub("s__NA",NA,BAC$Species)

SAMPLE_DATA$EJM_SOIL_NUMBER <- as.factor(SAMPLE_DATA$EJM_SOIL_NUMBER)

FUN_OTU<- FUN[,8:length(colnames(FUN))]
FUN_TAX <- FUN[,1:7]
BAC_OTU<- BAC[,8:length(colnames(BAC))]
BAC_TAX <- BAC[,1:7]

BAC_OTU <- BAC_OTU[rowSums(BAC_OTU)>1,]
FUN_OTU <- FUN_OTU[rowSums(FUN_OTU)>1,]

BAC_OTU <- BAC_OTU[,colSums(BAC_OTU)>1000]
FUN_OTU <- FUN_OTU[,colSums(FUN_OTU)>1000]

rownames(SAMPLE_DATA)<- SAMPLE_DATA$SEQUENCED_INV_NUMBER

SAMPLE_DATA_FIELD <- SAMPLE_DATA[SAMPLE_DATA$SOIL_SOURCE=="FIELD",]
```

Bacterial phyloseq

```{r}
OTU <- BAC_OTU[,colnames(BAC_OTU) %in% rownames(SAMPLE_DATA_FIELD)]

OTU <- OTU[rowSums(OTU)>1,]
OTU <- as.matrix(OTU)
TAX = BAC_TAX[rownames(OTU),]
samples = SAMPLE_DATA_FIELD[colnames(OTU),]

samples <- read.csv("../Data/output/sample_data_field_switch_samples_bac.csv")

sample_names(samples) <- samples$SEQUENCED_INV_NUMBER
rownames(samples) <- samples$SEQUENCED_INV_NUMBER


em_tse <- TreeSummarizedExperiment(assays = list(Count = OTU),
                                   rowData = TAX,
                                   colData = samples)
assayNames(em_tse) <- "counts" 
P.bac <- convertToPhyloseq(em_tse,assay.type = "counts")
P.bac = transform_sample_counts(P.bac, function(x) x / sum(x))
saveRDS(P.bac, "../Data/output/P.bac.fourfields.RDS")
```

Fungal phyloseq

```{r}
SAMPLE_DATA_FIELD <- SAMPLE_DATA[SAMPLE_DATA$SOIL_SOURCE=="FIELD",]
OTU <- FUN_OTU[,colnames(FUN_OTU) %in% rownames(SAMPLE_DATA_FIELD)]


OTU <- OTU[rowSums(OTU)>1,]
OTU <- as.matrix(OTU)
TAX = FUN_TAX[rownames(OTU),]
samples = SAMPLE_DATA_FIELD[colnames(OTU),]

samples <- read.csv("../Data/output/sample_data_field_switch_samples_fun.csv")


sample_names(samples) <- samples$SEQUENCED_INV_NUMBER
rownames(samples) <- samples$SEQUENCED_INV_NUMBER
em_tse <- TreeSummarizedExperiment(assays = list(Count = OTU),
                                   rowData = TAX,
                                   colData = samples)
assayNames(em_tse) <- "counts" 
P.fun <- convertToPhyloseq(em_tse,assay.type = "counts")
P.fun = transform_sample_counts(P.fun, function(x) x / sum(x))
saveRDS(P.fun, "../Data/output/P.fun.fourfields.RDS")
```


## PCOA

### Bacteria

```{r, warning=F}
data <- data.frame(t(otu_table(P.bac)))
data <- decostand(data, method='hellinger')
dist <- vegdist(data, method='bray')
pcoa_results <- cmdscale(dist, k = 2, eig = TRUE, add = TRUE) 
pcoa_positions <- as_tibble(pcoa_results$points)
colnames(pcoa_positions) <- c("PCoA1", "PCoA2") 
points <- scores(pcoa_results)
points<- as.data.frame(points)
colnames(points) <- c("PCoA1", "PCoA2") 
points$SEQUENCED_INV_NUMBER <- rownames(points)

eigenvalues <- pcoa_results$eig
percent_explained_PC1 <- round(100 * eigenvalues[1] / sum(eigenvalues), 1)
percent_explained_PC2 <- round(100 * eigenvalues[2] / sum(eigenvalues), 1)
d <- data.frame(sample_data(P.bac))
pcoa_data <- dplyr::inner_join(points,d, by = "SEQUENCED_INV_NUMBER")
pcoa_data$Year <- as.factor(pcoa_data$Year)

beta_bac <- ggplot(data = pcoa_data, aes(x=PCoA1,y=PCoA2)) +
  geom_point(aes(color = Farm, shape = Year, size = 5)) +
  scale_color_manual(values = c("#EC7063","#82E0AA","#5DADE2","#F7DC6F"))+
  labs(x = paste("PCoA 1 (", percent_explained_PC1, "%)", sep = ""),
    y = paste("PCoA 2 (", percent_explained_PC2, "%)", sep = "")) +
  theme_classic()+
  guides(size = "none", color = guide_legend(title = "Farm ID", theme = theme(legend.title = element_text(size = 12), legend.text = element_text(size = 12)),position = "right"),
         shape = guide_legend(title = "Year", theme = theme(legend.title = element_text(size = 12), legend.text = element_text(size = 12)),position = "right",ncol=1))+
  theme(legend.position = "none");beta_bac

```

### Fungi

```{r}
data <- data.frame(t(otu_table(P.fun)))
data <- decostand(data, method='hellinger')
dist <- vegdist(data, method='bray')
pcoa_results <- cmdscale(dist, k = 2, eig = TRUE, add = TRUE) 
pcoa_positions <- as_tibble(pcoa_results$points)
colnames(pcoa_positions) <- c("PCoA1", "PCoA2") 
points <- scores(pcoa_results)
points<- as.data.frame(points)
colnames(points) <- c("PCoA1", "PCoA2") 
points$SEQUENCED_INV_NUMBER <- rownames(points)

eigenvalues <- pcoa_results$eig
percent_explained_PC1 <- round(100 * eigenvalues[1] / sum(eigenvalues), 1)
percent_explained_PC2 <- round(100 * eigenvalues[2] / sum(eigenvalues), 1)
d <- data.frame(sample_data(P.fun))
pcoa_data <- dplyr::inner_join(points,d, by = "SEQUENCED_INV_NUMBER")
pcoa_data$Year <- as.factor(pcoa_data$Year)

beta_fun <- ggplot(data = pcoa_data, aes(x=PCoA1,y=PCoA2)) +
  geom_point(aes(color = Farm, shape = Year, size = 5)) +
  scale_color_manual(values = c("#EC7063","#82E0AA","#5DADE2","#F7DC6F"),labels = c("Crossroads", "Norwich Meadows", "Remembrance", "Turtle Tree"))+
  labs(x = paste("PCoA 1 (", percent_explained_PC1, "%)", sep = ""),
    y = paste("PCoA 2 (", percent_explained_PC2, "%)", sep = "")) +
  theme_classic()+
  guides(size = "none", color = guide_legend(title = "Farm ID", theme = theme(legend.title = element_text(size = 12), legend.text = element_text(size = 12)),position = "right"),
         shape = guide_legend(title = "Year", theme = theme(legend.title = element_text(size = 12), legend.text = element_text(size = 12)),position = "right",ncol=1));beta_fun
```

### Agglomerate plots

```{r}
BETA <- plot_grid(beta_bac, beta_fun, labels = c('B', 'C'), nrow = 1, ncol = 2, rel_widths = c(1, 1.35));BETA
ggsave(filename = "../Data/output/Field_sample_ordination.svg", plot = BETA,  width = 11, height = 4, device = "svg")
```

# Aphid bioassay

## Read in data

```{r}
data <- read_xlsx("../Data/HATCH_compost_bioassay_data.xlsx",sheet = "Formatted for R")

data <- data[data$F1==1&!is.na(data$F1),]
data <- data[data$NOTES!="SAMPLED ONE DAY EARLY"|is.na(data$NOTES),]
data <- data[data$PETIOLE_SNAP!="O"|is.na(data$PETIOLE_SNAP),]

data$combined <- paste(data$SLURRY,data$COMPOST)
data$combined = factor(data$combined, levels=c('HOAG NONE','HOAG COMP','CROSS NONE','CROSS COMP','REM NONE','REM COMP','NMF NONE', 'NMF COMP', 'TT NONE', 'TT COMP', 'SCROSS NONE', 'SCROSS COMP', 'SREM NONE','SREM COMP'))
data$F2 <- as.numeric(data$F2)
data$population <- data$F2 + 1

d <- data[data$SLURRY=="REM"|data$SLURRY=="CROSS"|data$SLURRY=="TT"|data$SLURRY=="NMF",]
d <- d[d$COMPOST=="NONE"|d$COMPOST=="COMP",]
d <- d[d$EXPERIMENT==1|d$EXPERIMENT==4,]
d$EXPERIMENT[d$EXPERIMENT==1] <- "2023"
d$EXPERIMENT[d$EXPERIMENT==4] <- "2024"
```

## Run model

```{r}
mod <- lm(population~COMPOST*SLURRY*EXPERIMENT, data=d)
par(mfrow=c(2,2))
plot(mod)
Anova(mod, Type = 3)
emmip(mod, COMPOST ~ SLURRY | EXPERIMENT, CIs = TRUE)

df <- as.data.frame(emmeans(mod, ~ SLURRY*EXPERIMENT))

res <- emmeans(mod, list(pairwise ~ SLURRY*EXPERIMENT), adjust = "fdr")
plot(res[[1]], horizontal=F)
groups <- data.frame(multcomp::cld(res[[1]], adjust='fdr', Letters=c("abcdefghijklmnopqrstuvwxy"), decreasing=T)) %>% 
  dplyr::select(., SLURRY, EXPERIMENT, .group)

df <- dplyr::inner_join(df,groups, by = c('SLURRY', 'EXPERIMENT'))

names(df)[names(df) == '.group'] <- 'letters'
df$letters <- gsub(' ','',df$letters)

dplyr::select(df, SLURRY, EXPERIMENT,letters)
d$letters <- ifelse(d$SLURRY == "REM" & d$EXPERIMENT == "2023", "c",  
                            ifelse(d$SLURRY == "REM" & d$EXPERIMENT == "2024", "ab",
                                   ifelse(d$SLURRY == "CROSS" & d$EXPERIMENT == "2023", "c",
                                          ifelse(d$SLURRY == "CROSS" & d$EXPERIMENT == "2024", "b",
                                                 ifelse(d$SLURRY == "NMF" & d$EXPERIMENT == "2023", "c",
                                                        ifelse(d$SLURRY == "NMF" & d$EXPERIMENT == "2024", "ab",
                                                               ifelse(d$SLURRY == "TT" & d$EXPERIMENT == "2023", "c",
                                                                      ifelse(d$SLURRY == "TT" & d$EXPERIMENT == "2024", "a",NA))))))))
```

## Generate figure

```{r, message = F}
aphid <- ggplot(data = d, aes(combined, population, fill = combined)) +
  geom_bar(stat = "summary", position = "dodge") +
  geom_jitter(width = 0.1, height = 0, color = "gray25", alpha = 0.6) +
  geom_errorbar(stat = "summary", position = "dodge", width=0.3, alpha=0.9, linewidth=1) +
geom_text(stat = "summary", aes(label = letters),position=position_dodge(0.9), vjust = -3, hjust = -0.2, size = 5)+
  scale_fill_manual(values = c("#EC7063","#B03A2E","#5DADE2","#21618C","#82E0AA","#239B56","#F7DC6F","#D4AC0D"))+
  scale_x_discrete(labels=c('Farm 1\nNo Compost','Farm 1\nCompost','Farm 2\nNo Compost','Farm 2\nCompost','Farm 3\nNo Compost','Farm 3\nCompost','Farm 4\nNo Compost','Farm 4\nCompost'))+
  facet_grid(rows = vars(EXPERIMENT)) +
  labs(x = "", y = "Aphid Population After 9 Days") + 
  ylim(0, 25) +
  theme_classic()+
  theme(legend.position = "none", strip.text = element_text(size = 12, color = "black"), 
        axis.title = element_text(size = 12, color = "black"),
        axis.text = element_text(size = 12, color = "black"));aphid
APHID<- plot_grid(aphid, nrow = 1,ncol = 1,labels = c('D'));APHID


ggsave(filename = "../Data/output/Figure_2D.svg", plot = APHID, width = 15, height = 8, device = "svg")
```

# Generates final figure

```{r}
fig2.top <- plot_grid(MAP, BETA, ncol = 2, nrow = 1, rel_widths = c(0.4,1));fig2.top
fig2 <- plot_grid(fig2.top, APHID, ncol=1,nrow=2, rel_heights = c(0.7,1));fig2
ggsave(filename = "../Data/output/Figure_2.svg", plot = fig2,  width = 12, height = 8, device = "svg")
```
