---
title: "Fig_1_Analysis"
author: "MD and EM"
date: "`r Sys.Date()`"
output: 
 html_document:
   toc: true
   toc_float: true
   editor_options:
     chunk_output_type: console
   chunck_output_type: console
---

Inputs: <br> Map<br>

-   Bloom_SiteLocations.csv <br>

ALPHA <br>

-   Formatted data on samples used in analysis.xlsx <br>

BETA <br>

-   Formatted data on samples used in analysis.xlsx
-   Bloom_otu_table_Bac.csv
-   Bloom_otu_table_Fun.csv
-   Bloom_taxonomy_table_Bac.csv
-   Bloom_taxonomy_table_Fun.csv
-   Bloom_sample_info.csv <br>

Outputs: <br>

-   figure 1
-   figure s1

# Load libraries

```{r, results = 'hide', message=FALSE}
library(tidyverse) # needed for ggplot2, map_data
library(cowplot) #plot_grid
library(readxl) # read_excel
library(car) #Anova
library(agricolae) # HSD.test
library(reshape2) #melt
library(phyloseq) #OTU_Table
library(vegan) # decostand
```

# Map

## Read in data

```{r}
site_locations <- read.csv("../Data/Bloom_SiteLocations.csv")
site_locations$Participant.ID <- as.factor(site_locations$Participant.ID)
```

## Structure df

```{r}
new_york <- map_data("county")
new_york<- new_york[new_york$region == "new york",]

compost_locations <- site_locations[
  site_locations$Participant.ID =="132" |
    site_locations$Participant.ID =="118"|
    site_locations$Participant.ID =="156"|
    site_locations$Participant.ID =="172"|
    site_locations$Participant.ID =="21"|
    site_locations$Participant.ID =="213"|
    site_locations$Participant.ID =="223"|
    site_locations$Participant.ID =="226"|
    site_locations$Participant.ID =="232"|
    site_locations$Participant.ID =="244",]
```

## Generate figure

```{r}
map <- ggplot() + 
  geom_polygon( data=new_york, aes(x=long, y=lat, group = group),  fill="lightblue" ) +
  geom_point(data = compost_locations, size = 5, aes(x=Longitude, y=Latitude, color = Participant.ID)) +
  theme_classic()+
  labs(x= "Latitude", y = "Longitude")+
  theme_classic()+
  theme(legend.position = "none", strip.text = element_text(size = 15, color = "black"), 
        axis.title = element_text(size = 15, color = "black"),
        axis.text = element_text(size = 12, color = "black"));map

MAP <- plot_grid(map, labels = "A", label_size = 20);MAP
```

# Alpha diveristy

## Read in data

```{r}
data <- read_excel("../Data/Formatted data on samples used in analysis.xlsx",sheet="Data")
data$combined <- paste(data$`Participant Number`,data$B.E_10_Compost)
data$`Participant Number`<- as.factor(data$`Participant Number`)
data$combined <- as.factor(data$combined)
data$B.E_10_Compost <- as.factor(data$B.E_10_Compost)

sha.long <- dplyr::select(data, c("Extracted ID", "Cleaned ID", "Sequenced ID", "Participant Number", "bac.shan", "fun.shan", "B.E_10_Compost")) %>% 
  pivot_longer(., cols = c(bac.shan,fun.shan), names_to = "Taxa", values_to = "Shannon")

otu.long <- dplyr::select(data, c("Extracted ID", "Cleaned ID", "Sequenced ID", "Participant Number", "bac.rich", "fun.rich", "B.E_10_Compost")) %>% 
  pivot_longer(., cols = c(bac.rich,fun.rich), names_to = "Taxa", values_to = "OTU_Rich")

sim.long <- dplyr::select(data, c("Extracted ID", "Cleaned ID", "Sequenced ID", "Participant Number", "bac.sim", "fun.sim", "B.E_10_Compost")) %>% 
  pivot_longer(., cols = c(bac.sim,fun.sim), names_to = "Taxa", values_to = "Simp")

par(mfrow=c(2,2))
options(contrasts = c("contr.sum", "contr.poly"))
```

## Diversity models

### Shannon model

```{r}
mod <- lm(Shannon ~ B.E_10_Compost*Taxa, data = sha.long)
plot(mod)

Anova(mod, Type = 3)
t <- HSD.test(mod, "Taxa");t
```

### OTU Richness

```{r, message=F}
mod <- lm(OTU_Rich ~ B.E_10_Compost*Taxa, data = otu.long)
plot(mod)

Anova(mod, Type = 3)
t <- HSD.test(mod, "Taxa");t
```

### Simpsons

```{r}
mod <- lm(Simp ~ B.E_10_Compost*Taxa, data = sim.long)
plot(mod)

Anova(mod, Type = 3)
t <- HSD.test(mod, "Taxa");t
```

## Generates figures

### Bacteria

OTU Richness

```{r, message=F}
bac_rich <- ggplot(data, aes(B.E_10_Compost,bac.rich)) +
  geom_bar(stat = "summary", position = "dodge", fill = "lightblue") +
  geom_jitter(width = 0.1, height = 0, aes(color = `Participant Number`), size = 3) +
  geom_errorbar(stat = "summary", position = "dodge", width=0.4, alpha=0.9, linewidth=0.5) +
  xlab(label = "") +
  ylab(label = "Bacterial OTU Richness") +
  scale_x_discrete(labels=c('No Compost','Compost'))+
  theme_classic()+
  theme(legend.position = "none", strip.text = element_text(size = 15, color = "black"), 
        axis.title = element_text(size = 15, color = "black"),
        axis.text = element_text(size = 12, color = "black"));bac_rich
```

Shannon

```{r, message=F}
bac_shan <- ggplot(data, aes(B.E_10_Compost,bac.shan)) +
  geom_bar(stat = "summary", position = "dodge", fill = "lightblue") +
  geom_jitter(width = 0.1, height = 0, aes(color = `Participant Number`), size = 3) +
  geom_errorbar(stat = "summary", position = "dodge", width=0.4, alpha=0.9, linewidth=0.5) +
  xlab(label = "") +
  ylab(label = "Bacterial Shannon Diversity") +
  scale_x_discrete(labels=c('No Compost','Compost'))+
  theme_classic()+
  theme(legend.position = "none", strip.text = element_text(size = 15, color = "black"), 
        axis.title = element_text(size = 15, color = "black"),
        axis.text = element_text(size = 12, color = "black"));bac_shan
```

Simpsons

```{r, message=F}
bac_sim <- ggplot(data, aes(B.E_10_Compost,bac.sim)) +
  geom_bar(stat = "summary", position = "dodge", fill = "lightblue") +
  geom_jitter(width = 0.1, height = 0, aes(color = `Participant Number`), size = 3) +
  geom_errorbar(stat = "summary", position = "dodge", width=0.4, alpha=0.9, linewidth=0.5) +
  xlab(label = "") +
  ylab(label = "Bacterial Simpsons Diversity") +
  scale_x_discrete(labels=c('No Compost','Compost'))+
  theme_classic()+
  theme(legend.position = "none", strip.text = element_text(size = 15, color = "black"), 
        axis.title = element_text(size = 15, color = "black"),
        axis.text = element_text(size = 12, color = "black"));bac_sim
```

### Fungi

OTU Richness

```{r, message=F}
fun_rich <- ggplot(data, aes(B.E_10_Compost,fun.rich)) +
  geom_bar(stat = "summary", position = "dodge", fill = "lightblue") +
  geom_jitter(width = 0.1, height = 0, aes(color = `Participant Number`), size = 3) +
  geom_errorbar(stat = "summary", position = "dodge", width=0.4, alpha=0.9, linewidth=0.5) +
  xlab(label = "") +
  ylab(label = "Fungal OTU Richness") +
  scale_x_discrete(labels=c('No Compost','Compost'))+
  theme_classic()+
  theme(legend.position = "none", strip.text = element_text(size = 15, color = "black"), 
        axis.title = element_text(size = 15, color = "black"),
        axis.text = element_text(size = 12, color = "black"));fun_rich
```

Shannons

```{r, message=F}
fun_shan <- ggplot(data, aes(B.E_10_Compost,fun.shan)) +
  geom_bar(stat = "summary", position = "dodge", fill = "lightblue") +
  geom_jitter(width = 0.1, height = 0, aes(color = `Participant Number`), size = 3) +
  geom_errorbar(stat = "summary", position = "dodge", width=0.4, alpha=0.9, linewidth=0.5) +
  xlab(label = "") +
  ylab(label = "Fungal Shannon Diversity") +
  scale_x_discrete(labels=c('No Compost','Compost'))+
  theme_classic()+
  theme(legend.position = "none", strip.text = element_text(size = 15, color = "black"), 
        axis.title = element_text(size = 15, color = "black"),
        axis.text = element_text(size = 12, color = "black"));fun_shan
```

Simpsons

```{r, message=F}
fun_sim <- ggplot(data, aes(B.E_10_Compost,fun.sim)) +
  geom_bar(stat = "summary", position = "dodge", fill = "lightblue") +
  geom_jitter(width = 0.1, height = 0, aes(color = `Participant Number`), size = 3) +
  geom_errorbar(stat = "summary", position = "dodge", width=0.4, alpha=0.9, linewidth=0.5) +
  xlab(label = "") +
  ylab(label = "Fungal Simpsons Diversity") +
  scale_x_discrete(labels=c('No Compost','Compost'))+
  theme_classic()+
  theme(legend.position = "none", strip.text = element_text(size = 15, color = "black"), 
        axis.title = element_text(size = 15, color = "black"),
        axis.text = element_text(size = 12, color = "black"));fun_sim
```

### Generates agglomerated figures

Supplemental figure

```{r, message=F}
sup <- plot_grid(bac_rich,bac_shan,bac_sim,fun_rich,fun_shan,fun_sim, nrow = 2,ncol = 3,labels = "AUTO",label_size = 20);sup
ggsave(filename = "../Data/output/Supplemental_figure_1.svg", plot = sup, width = 11, height = 10, device = "svg")
```

Main text figure

```{r, message=F}
ALPHA <- plot_grid(bac_shan,fun_shan,nrow = 2, ncol = 1, labels = c("B","C"),label_size = 20);ALPHA
```

# Ordination

## Read in data

```{r}
data <- read_excel("../Data/Formatted data on samples used in analysis.xlsx",sheet="Data")
data$combined <- paste(data$`Participant Number`,data$B.E_10_Compost)
data$`Participant Number`<- as.factor(data$`Participant Number`)
data$combined <- as.factor(data$combined)
data$B.E_10_Compost <- as.factor(data$B.E_10_Compost)


bacterial_otu_mat <- read.csv("../Data/Bloom_otu_table_Bac.csv")
fungal_otu_mat <- read.csv("../Data/Bloom_otu_table_Fun.csv")
bacterial_tax_mat <- read.csv("../Data/Bloom_taxonomy_table_Bac.csv")
fungal_tax_mat <- read.csv("../Data/Bloom_taxonomy_table_Fun.csv")
comb_samples_df <- read.csv("../Data/Bloom_sample_info.csv")

###converts relevant column to rownames to each object
bacterial_otu_mat <- bacterial_otu_mat %>%
  tibble::column_to_rownames("X")
fungal_otu_mat <- fungal_otu_mat %>%
  tibble::column_to_rownames("X")
bacterial_tax_mat <- bacterial_tax_mat %>%
  tibble::column_to_rownames("X")
fungal_tax_mat <- fungal_tax_mat %>%
  tibble::column_to_rownames("X")
comb_samples_df <- comb_samples_df %>%
  tibble::column_to_rownames("IMR.Label")

###adds taxonomic prefixes to bacterial tax table
bacterial_tax_mat$Kingdom<-sub("^","k__",bacterial_tax_mat$Kingdom)
bacterial_tax_mat$Phylum<-sub("^","p__",bacterial_tax_mat$Phylum)
bacterial_tax_mat$Class<-sub("^","c__",bacterial_tax_mat$Class)
bacterial_tax_mat$Order<-sub("^","o__",bacterial_tax_mat$Order)
bacterial_tax_mat$Family<-sub("^","f__",bacterial_tax_mat$Family)
bacterial_tax_mat$Genus<-sub("^","g__",bacterial_tax_mat$Genus)
bacterial_tax_mat$Species<-sub("^","s__",bacterial_tax_mat$Species)
```

## Formats df for phyloseq creation

```{r, message=F}
###makes a new df of OTU# and the deepest taxonomy alignment for each bacterial and fungal OTU 
bac_combined_tax <- if_else(bacterial_tax_mat$Species == "_","test_true",bacterial_tax_mat$Species, missing = bacterial_tax_mat$Genus)
bac_combined_tax <- if_else(bac_combined_tax == "_","test_true",bac_combined_tax, missing = bacterial_tax_mat$Family)
bac_combined_tax <- if_else(bac_combined_tax == "_","test_true",bac_combined_tax, missing = bacterial_tax_mat$Order)
bac_combined_tax <- if_else(bac_combined_tax == "_","test_true",bac_combined_tax, missing = bacterial_tax_mat$Class)
bac_combined_tax <- if_else(bac_combined_tax == "_","test_true",bac_combined_tax, missing = bacterial_tax_mat$Phylum)
bac_combined_tax <- if_else(bac_combined_tax == "_","test_true",bac_combined_tax, missing = bacterial_tax_mat$Kingdom)
bac_combined_tax <- melt(data.frame(rownames(bacterial_tax_mat),bac_combined_tax))
colnames(bac_combined_tax) <- c("OTU", "taxon")
bac_combined_tax$sequence <- bacterial_tax_mat$sequence

fun_combined_tax <- if_else(fungal_tax_mat$Species == "_","test_true",fungal_tax_mat$Species, missing = fungal_tax_mat$Genus)
fun_combined_tax <- if_else(fun_combined_tax == "_","test_true",fun_combined_tax, missing = fungal_tax_mat$Family)
fun_combined_tax <- if_else(fun_combined_tax == "_","test_true",fun_combined_tax, missing = fungal_tax_mat$Order)
fun_combined_tax <- if_else(fun_combined_tax == "_","test_true",fun_combined_tax, missing = fungal_tax_mat$Class)
fun_combined_tax <- if_else(fun_combined_tax == "_","test_true",fun_combined_tax, missing = fungal_tax_mat$Phylum)
fun_combined_tax <- if_else(fun_combined_tax == "_","test_true",fun_combined_tax, missing = fungal_tax_mat$Kingdom)
fun_combined_tax <- melt(data.frame(rownames(fungal_tax_mat),fun_combined_tax))
colnames(fun_combined_tax) <- c("OTU", "taxon")
fun_combined_tax$sequence <- fungal_tax_mat$sequence

###converts Participant.ID, composting, and relevant biopesticides to factors
comb_samples_df$Participant.ID <- as.factor(comb_samples_df$Participant.ID)
comb_samples_df$B.E_10_Compost <- as.factor(comb_samples_df$B.E_10_Compost)

###adds a column to comb_samples_df to indicate unique sample ID
comb_samples_df$IMR_label <- rownames(comb_samples_df)

###converts otu tables and tax tables to matrices
bacterial_otu_mat <- as.matrix(bacterial_otu_mat)
fungal_otu_mat <- as.matrix(fungal_otu_mat)
bacterial_tax_mat <- as.matrix(bacterial_tax_mat)
fungal_tax_mat <- as.matrix(fungal_tax_mat)

###converts all bacterial tables to single bacterial_phyloseq object
OTU = otu_table(bacterial_otu_mat, taxa_are_rows = TRUE)
TAX = tax_table(bacterial_tax_mat)
samples = sample_data(comb_samples_df)

bacterial_phyloseq <- phyloseq(OTU, TAX, samples)
bacterial_phyloseq

###converts all fungal tables to single fungal_phyloseq object
OTU = otu_table(fungal_otu_mat, taxa_are_rows = TRUE)
TAX = tax_table(fungal_tax_mat)
samples = sample_data(comb_samples_df)

fungal_phyloseq <- phyloseq(OTU, TAX, samples)
fungal_phyloseq


###makes subset of bacterial and fungal samples for 20 paired composted/uncomposted sites
bac_compost_phyloseq <- subset_samples(bacterial_phyloseq, 
                                       Participant.ID =="132" |
                                         Participant.ID =="118"|
                                         Participant.ID =="156"|
                                         Participant.ID =="172"| 
                                         Participant.ID =="21"| 
                                         Participant.ID =="213"| 
                                         Participant.ID =="223"| 
                                         Participant.ID =="226"| 
                                         Participant.ID =="232"| 
                                         Participant.ID =="244")
sample_names(bac_compost_phyloseq)
fun_compost_phyloseq <- subset_samples(fungal_phyloseq, 
                                       Participant.ID =="132" | 
                                         Participant.ID =="118"| 
                                         Participant.ID =="156"| 
                                         Participant.ID =="172"| 
                                         Participant.ID =="21"| 
                                         Participant.ID =="213"| 
                                         Participant.ID =="223"| 
                                         Participant.ID =="226"| 
                                         Participant.ID =="232"| 
                                         Participant.ID =="244")
sample_names(fun_compost_phyloseq)
```

## Generates PCoA for Bacteria

### Creates df for PCoA

```{r}
P.bac.fig1.rel = transform_sample_counts(bac_compost_phyloseq, function(x) x / sum(x))

d <- data.frame(sample_data(P.bac.fig1.rel))
data <- data.frame(t(otu_table(P.bac.fig1.rel)))
data <- decostand(data, method='hellinger')
dist <- vegdist(data, method='bray')

pcoa_results <- cmdscale(dist, k = 2, eig = TRUE, add = TRUE) # `add=TRUE` handles negative eigenvalues
pcoa_positions <- as_tibble(pcoa_results$points)
colnames(pcoa_positions) <- c("PCoA1", "PCoA2") 
points <- scores(pcoa_results)
points<- as.data.frame(points)
colnames(points) <- c("PCoA1", "PCoA2") 
points$IMR_label <- rownames(points)

eigenvalues <- pcoa_results$eig
percent_explained_PC1 <- round(100 * eigenvalues[1] / sum(eigenvalues), 1)
percent_explained_PC2 <- round(100 * eigenvalues[2] / sum(eigenvalues), 1)

d <- data.frame(sample_data(P.bac.fig1.rel))
pcoa_data <- dplyr::inner_join(points,d, by = "IMR_label")

beta_bac <- ggplot(data = pcoa_data, aes(x=PCoA1,y=PCoA2)) +
  geom_point(aes(color = Participant.ID,shape = B.E_10_Compost, size = 5))+
  scale_shape_discrete(labels = c("No", "Yes")) +
  labs(x = paste("PCoA 1 (", percent_explained_PC1, "%)", sep = ""),
    y = paste("PCoA 2 (", percent_explained_PC2, "%)", sep = "")) +
  theme_classic()+
  guides(size = "none", color = guide_legend(title = "Farm ID", theme = theme(legend.title = element_text(size = 12), legend.text = element_text(size = 12)),position = "bottom"),
         shape = guide_legend(title = "Compost", theme = theme(legend.title = element_text(size = 12), legend.text = element_text(size = 12)),position = "bottom",ncol=1));beta_bac
```

## Generates PCoA for Fungi

```{r}
P.fun.fig1.rel = transform_sample_counts(fun_compost_phyloseq, function(x) x / sum(x))

d <- data.frame(sample_data(P.fun.fig1.rel))
data <- data.frame(t(otu_table(P.fun.fig1.rel)))
data <- decostand(data, method='hellinger')
dist <- vegdist(data, method='bray')

pcoa_results <- cmdscale(dist, k = 2, eig = TRUE, add = TRUE) # `add=TRUE` handles negative eigenvalues

pcoa_positions <- as_tibble(pcoa_results$points)
colnames(pcoa_positions) <- c("PCoA1", "PCoA2") 
points <- scores(pcoa_results)
points<- as.data.frame(points)
colnames(points) <- c("PCoA1", "PCoA2") 
points$IMR_label <- rownames(points)

eigenvalues <- pcoa_results$eig
percent_explained_PC1 <- round(100 * eigenvalues[1] / sum(eigenvalues), 1)
percent_explained_PC2 <- round(100 * eigenvalues[2] / sum(eigenvalues), 1)

d <- data.frame(sample_data(P.fun.fig1.rel))
pcoa_data <- dplyr::inner_join(points,d, by = "IMR_label")

beta_fun <- ggplot(data = pcoa_data, aes(x=PCoA1,y=PCoA2)) +
  geom_point(aes(color = Participant.ID,shape = B.E_10_Compost, size = 5))+
  scale_shape_discrete(labels = c("No", "Yes")) +
  theme_classic()+
  labs(x = paste("PCoA 1 (", percent_explained_PC1, "%)", sep = ""),
    y = paste("PCoA 2 (", percent_explained_PC2, "%)", sep = ""))+
  theme(legend.position = "none");beta_fun
```

## Agglomerates Bacteria and Fungi PCoAs

```{r}
BETA <- plot_grid(beta_bac, beta_fun, ncol = 1, nrow = 2,labels = c("D","E"),label_size = 20, rel_heights = c(1.2,1));BETA
```

# Generate Figure 1

Agglomerates `MAP`, `ALPHA`, `BETA`

```{r}
fig_1<- plot_grid(MAP,ALPHA,BETA, nrow = 1,ncol = 3, rel_widths = c(2, 1, 1.5));fig_1
ggsave(filename = "../Data/output/Figure_1.svg", plot = fig_1,  width = 18, height = 8, device = "svg")
```
