---
title: "Untitled"
author: "MD"
date: "`r Sys.Date()`"
output: 
 html_document:
   toc: true
   toc_float: true
   editor_options:
   chunck_output_type: console
---
Inputs: <br>

-   EJM-CLS-22F_pct98.csv
-   EJM-CLS-22B_pct98.csv
-   Sequencing_data_formatted_for_analysis(01.28.2026).xlsx
-   Bloom_otu_table_Bac.csv
-   Bloom_otu_table_Fun.csv
-   Bloom_taxonomy_table_Bac.csv
-   Bloom_taxonomy_table_Fun.csv
-   Bloom_sample_info.csv
 

Outputs: <br>

-   figure 6


# Load libraries
```{r, results = 'hide', message=FALSE}
knitr::opts_chunk$set(out.width = "100%")
library(readxl)
library(ANCOMBC)
library(tidyverse)
library(reshape2)
library(phyloseq)
library(cowplot)
```

# Lab trial

## Read in data
```{r, message=F, warning=F}
FUN <- read.csv("../Data/EJM-CLS-22F_pct98.csv")
BAC <- read.csv("../Data/EJM-CLS-22B_pct98.csv")
BAC <- BAC[!is.na(BAC$Kingdom),]
FUN <- FUN[FUN$Kingdom =="k__Fungi",]
FUN <- FUN[!is.na(FUN$Kingdom),]

SAMPLE_DATA <- as.data.frame(read_xlsx("../Data/Sequencing_data_formatted_for_analysis(01.28.2026).xlsx", sheet = "Soil_sample_info"))
```

## Create phyloseq
```{r, message=F, warning=F}
rownames(FUN) <- FUN$X
FUN <- FUN[,-1]
rownames(BAC) <- BAC$X
BAC <- BAC[,-1]

BAC$Kingdom<-sub("^","k__",BAC$Kingdom)
BAC$Kingdom<-sub("k__NA",NA,BAC$Kingdom)
BAC$Phylum<-sub("^","p__",BAC$Phylum)
BAC$Phylum<-sub("p__NA",NA,BAC$Phylum)
BAC$Class<-sub("^","c__",BAC$Class)
BAC$Class<-sub("c__NA",NA,BAC$Class)
BAC$Order<-sub("^","o__",BAC$Order)
BAC$Order<-sub("o__NA",NA,BAC$Order)
BAC$Family<-sub("^","f__",BAC$Family)
BAC$Family<-sub("f__NA",NA,BAC$Family)
BAC$Genus<-sub("^","g__",BAC$Genus)
BAC$Genus<-sub("g__NA",NA,BAC$Genus)
BAC$Species<-sub("^","s__",BAC$Species)
BAC$Species<-sub("s__NA",NA,BAC$Species)

SAMPLE_DATA$EJM_SOIL_NUMBER <- as.factor(SAMPLE_DATA$EJM_SOIL_NUMBER)


FUN_OTU<- FUN[,8:length(colnames(FUN))]
FUN_TAX <- FUN[,1:7]
BAC_OTU<- BAC[,8:length(colnames(BAC))]
BAC_TAX <- BAC[,1:7]

BAC_TAX <- as.data.frame(BAC_TAX)
FUN_TAX <- as.data.frame(FUN_TAX)

BAC_TAX$Species[grep('incertae', BAC_TAX$Species, ignore.case = TRUE)] <- NA
BAC_TAX$Species[grep('unknown', BAC_TAX$Species, ignore.case = TRUE)] <- NA
BAC_TAX$Genus[grep('incertae', BAC_TAX$Genus, ignore.case = TRUE)] <- NA
BAC_TAX$Genus[grep('unknown', BAC_TAX$Genus, ignore.case = TRUE)] <- NA
BAC_TAX$Family[grep('incertae', BAC_TAX$Family, ignore.case = TRUE)] <- NA
BAC_TAX$Family[grep('unknown', BAC_TAX$Family, ignore.case = TRUE)] <- NA
FUN_TAX$Species[grep('incertae', FUN_TAX$Species, ignore.case = TRUE)] <- NA
FUN_TAX$Species[grep('unknown', FUN_TAX$Species, ignore.case = TRUE)] <- NA
FUN_TAX$Genus[grep('incertae', FUN_TAX$Genus, ignore.case = TRUE)] <- NA
FUN_TAX$Genus[grep('unknown', FUN_TAX$Genus, ignore.case = TRUE)] <- NA
FUN_TAX$Family[grep('incertae', FUN_TAX$Family, ignore.case = TRUE)] <- NA
FUN_TAX$Family[grep('unknown', FUN_TAX$Family, ignore.case = TRUE)] <- NA

BAC_TAX$Species <- if_else(is.na(BAC_TAX$Species),NA,paste(BAC_TAX$Genus,"_",BAC_TAX$Species,sep = ""))
FUN_TAX$Species <- if_else(is.na(FUN_TAX$Species),NA,paste("s__",substr(FUN_TAX$Genus, 4, nchar(FUN_TAX$Genus)),"_",substr(FUN_TAX$Species, 4, nchar(FUN_TAX$Species)),sep = ""))

bac_combined_tax <- if_else(BAC_TAX$Species == "_","test_true",BAC_TAX$Species, missing = BAC_TAX$Genus)
bac_combined_tax <- if_else(bac_combined_tax == "_","test_true",bac_combined_tax, missing = BAC_TAX$Family)
bac_combined_tax <- if_else(bac_combined_tax == "_","test_true",bac_combined_tax, missing = BAC_TAX$Order)
bac_combined_tax <- if_else(bac_combined_tax == "_","test_true",bac_combined_tax, missing = BAC_TAX$Class)
bac_combined_tax <- if_else(bac_combined_tax == "_","test_true",bac_combined_tax, missing = BAC_TAX$Phylum)
bac_combined_tax <- if_else(bac_combined_tax == "_","test_true",bac_combined_tax, missing = BAC_TAX$Kingdom)
bac_combined_tax <- melt(data.frame(rownames(BAC_TAX),bac_combined_tax))
colnames(bac_combined_tax) <- c("OTU", "taxon")

fun_combined_tax <- if_else(FUN_TAX$Species == "_","test_true",FUN_TAX$Species, missing = FUN_TAX$Genus)
fun_combined_tax <- if_else(fun_combined_tax == "_","test_true",fun_combined_tax, missing = FUN_TAX$Family)
fun_combined_tax <- if_else(fun_combined_tax == "_","test_true",fun_combined_tax, missing = FUN_TAX$Order)
fun_combined_tax <- if_else(fun_combined_tax == "_","test_true",fun_combined_tax, missing = FUN_TAX$Class)
fun_combined_tax <- if_else(fun_combined_tax == "_","test_true",fun_combined_tax, missing = FUN_TAX$Phylum)
fun_combined_tax <- if_else(fun_combined_tax == "_","test_true",fun_combined_tax, missing = FUN_TAX$Kingdom)
fun_combined_tax <- melt(data.frame(rownames(FUN_TAX),fun_combined_tax))
colnames(fun_combined_tax) <- c("OTU", "taxon")

SAMPLE_DATA$EJM_SOIL_NUMBER <- as.factor(SAMPLE_DATA$EJM_SOIL_NUMBER)
SAMPLE_DATA$Compost <- factor(SAMPLE_DATA$Compost, levels = c("NONE","COMP"))

SAMPLE_DATA_LAB <- SAMPLE_DATA[SAMPLE_DATA$SOIL_SOURCE=="LAB",]
SAMPLE_DATA_LAB <- SAMPLE_DATA_LAB[!is.na(SAMPLE_DATA_LAB$Collection_Date),]
SAMPLE_DATA_LAB <- SAMPLE_DATA_LAB[SAMPLE_DATA_LAB$Collection_Date!="45777",]
rownames(SAMPLE_DATA_LAB) <- SAMPLE_DATA_LAB$SEQUENCED_INV_NUMBER

FUN_OTU <- FUN_OTU[,colnames(FUN_OTU)%in%SAMPLE_DATA_LAB$SEQUENCED_INV_NUMBER]
BAC_OTU <- BAC_OTU[,colnames(BAC_OTU)%in%SAMPLE_DATA_LAB$SEQUENCED_INV_NUMBER]

BAC_OTU <- BAC_OTU[rowSums(BAC_OTU)>1,]
FUN_OTU <- FUN_OTU[rowSums(FUN_OTU)>1,]


FUN_OTU<- as.matrix(FUN_OTU)
FUN_TAX <- as.matrix(FUN_TAX)
BAC_OTU<- as.matrix(BAC_OTU)
BAC_TAX <- as.matrix(BAC_TAX)

OTU = otu_table(FUN_OTU, taxa_are_rows = TRUE)
TAX = tax_table(FUN_TAX)
samples = sample_data(SAMPLE_DATA_LAB)
sample_names(SAMPLE_DATA_LAB) <- SAMPLE_DATA_LAB$SEQUENCED_INV_NUMBER
FUN_phyloseq <- phyloseq(OTU, TAX, samples)

OTU = otu_table(BAC_OTU, taxa_are_rows = TRUE)
TAX = tax_table(BAC_TAX)
samples = sample_data(SAMPLE_DATA_LAB)
sample_names(SAMPLE_DATA_LAB) <- SAMPLE_DATA_LAB$SEQUENCED_INV_NUMBER
BAC_phyloseq <- phyloseq(OTU, TAX, samples)


lab_bac_compost_tse = mia::convertFromPhyloseq(BAC_phyloseq)
lab_fun_compost_tse = mia::convertFromPhyloseq(FUN_phyloseq)
```

## Runs ancombc
```{r, message=F, warning=F}
set.seed(123)
Sys.time()
compost_ancom_bac_SPE = ancombc2(data = lab_bac_compost_tse, assay_name = "counts",tax_level = "Species",
                                 fix_formula = "Compost*Slurry",
                                 p_adj_method = "holm", pseudo_sens = TRUE,
                                 prv_cut = 0.10, lib_cut = 0, s0_perc = 0.05,
                                 group = NULL, struc_zero = FALSE, neg_lb = FALSE,
                                 alpha = 0.05, n_cl = 4, verbose = TRUE,
                                 global = FALSE, pairwise = FALSE, dunnet = FALSE, trend = FALSE,
                                 iter_control = list(tol = 1e-2, max_iter = 20, verbose = TRUE),
                                 em_control = list(tol = 1e-5, max_iter = 100),
                                 lme_control = lme4::lmerControl(),
                                 mdfdr_control = list(fwer_ctrl_method = "holm", B = 100),
                                 trend_control = NULL)
Sys.time()
#for fungi
set.seed(123)
Sys.time()
compost_ancom_fun_SPE = ancombc2(data = lab_fun_compost_tse, assay_name = "counts",tax_level = "Species",
                                 fix_formula = "Compost*Slurry",
                                 p_adj_method = "holm", pseudo_sens = TRUE,
                                 prv_cut = 0.10, lib_cut = 0, s0_perc = 0.05,
                                 group = NULL, struc_zero = FALSE, neg_lb = FALSE,
                                 alpha = 0.05, n_cl = 4, verbose = TRUE,
                                 global = FALSE, pairwise = FALSE, dunnet = FALSE, trend = FALSE,
                                 iter_control = list(tol = 1e-2, max_iter = 20, verbose = TRUE),
                                 em_control = list(tol = 1e-5, max_iter = 100),
                                 lme_control = lme4::lmerControl(),
                                 mdfdr_control = list(fwer_ctrl_method = "holm", B = 100),
                                 trend_control = NULL)
Sys.time()
```

### Save files 
```{r, message=F, warning=F}
saveRDS(compost_ancom_bac_SPE, file = '../Data/output/Long_read_compost_ancom_bac_SPE.rds', ascii = FALSE, version = NULL,
        compress = TRUE, refhook = NULL)
saveRDS(compost_ancom_fun_SPE, file = '../Data/output/Long_read_compost_ancom_fun_SPE.rds', ascii = FALSE, version = NULL,
        compress = TRUE, refhook = NULL)


compost_ancom_bac_SPE <- readRDS(file = '../Data/output/Long_read_compost_ancom_bac_SPE.rds', refhook = NULL)
compost_ancom_fun_SPE <- readRDS(file = '../Data/output/Long_read_compost_ancom_fun_SPE.rds', refhook = NULL)
```

## Generates figure for bacteria

### Prep df for figure
```{r, message=F, warning=F}
res_prim = compost_ancom_bac_SPE$res

df_fig_compost_volcano = res_prim %>%
  dplyr::filter(p_CompostCOMP <1) %>%
  dplyr::arrange(desc(lfc_CompostCOMP)) %>%
  dplyr::filter(passed_ss_CompostCOMP == 1) %>%
  dplyr::mutate(direct = ifelse(lfc_CompostCOMP > 0, "Positive LFC", "Negative LFC"))
df_fig_compost_volcano$taxon = factor(df_fig_compost_volcano$taxon, levels = df_fig_compost_volcano$taxon)
df_fig_compost_volcano$direct = factor(df_fig_compost_volcano$direct,
                               levels = c("Positive LFC", "Negative LFC"))


df_fig_compost_volcano$sig <- ifelse(df_fig_compost_volcano$p_CompostCOMP>0.01, "NS",
                                       ifelse(df_fig_compost_volcano$p_CompostCOMP==0.01, "NS",
                                              ifelse(df_fig_compost_volcano$p_CompostCOMP<0.01, "sig",NA)))
df_fig_compost_volcano$sig.stage <- paste(df_fig_compost_volcano$direct,df_fig_compost_volcano$sig)


df_fig_compost_volcano$colors <- ifelse(df_fig_compost_volcano$sig=="NS", "NS",
                        ifelse(df_fig_compost_volcano$sig.stage=='Positive LFC sig',"Compost",
                               ifelse(df_fig_compost_volcano$sig.stage=='Negative LFC sig','No Compost',NA)))
df_fig_compost_volcano<- arrange(df_fig_compost_volcano, -log(p_CompostCOMP))
```

### Makes figure
```{r}
p.bac.lab<-ggplot(data=df_fig_compost_volcano)+
  geom_point(alpha=0.85,mapping= aes(
    fill=colors, x=lfc_CompostCOMP,y=-log(p_CompostCOMP)),
    size=5,color='black',pch=21) +
  scale_fill_manual(values = c('chocolate4','cyan4','gray50'))+
  theme_classic()+
  labs(y=expression("-log"["10"] ~ "p" ["adj"]), 
       x=expression("log"["2"]~"Fold Change"))+
  theme(axis.text=element_text(size=10, colour = "black"), 
        axis.line = element_line(colour = "black"),
        legend.position="none");p.bac.lab
ggsave(filename = "../Data/output/volcano.bac.lab.svg", plot = p.bac.lab,
       width = 8, height = 8, device = "svg")
```

## Generates figure for fungi 

### Makes df for figure
```{r, message=F, warning=F}
res_prim = compost_ancom_fun_SPE$res

df_fig_compost_volcano = res_prim %>%
  dplyr::filter(p_CompostCOMP <1) %>%
  dplyr::arrange(desc(lfc_CompostCOMP)) %>%
  dplyr::filter(passed_ss_CompostCOMP == 1) %>%
  dplyr::mutate(direct = ifelse(lfc_CompostCOMP > 0, "Positive LFC", "Negative LFC"))
df_fig_compost_volcano$taxon = factor(df_fig_compost_volcano$taxon, levels = df_fig_compost_volcano$taxon)
df_fig_compost_volcano$direct = factor(df_fig_compost_volcano$direct,
                               levels = c("Positive LFC", "Negative LFC"))


df_fig_compost_volcano$sig <- ifelse(df_fig_compost_volcano$p_CompostCOMP>0.01, "NS",
                                       ifelse(df_fig_compost_volcano$p_CompostCOMP==0.01, "NS",
                                              ifelse(df_fig_compost_volcano$p_CompostCOMP<0.01, "sig",NA)))
df_fig_compost_volcano$sig.stage <- paste(df_fig_compost_volcano$direct,df_fig_compost_volcano$sig)


df_fig_compost_volcano$colors <- ifelse(df_fig_compost_volcano$sig=="NS", "NS",
                        ifelse(df_fig_compost_volcano$sig.stage=='Positive LFC sig',"Compost",
                               ifelse(df_fig_compost_volcano$sig.stage=='Negative LFC sig','No Compost',NA)))
df_fig_compost_volcano<- arrange(df_fig_compost_volcano, -log(p_CompostCOMP))
```

### makes fig
```{r, message=F, warning=F}
p.fun.lab<-ggplot(data=df_fig_compost_volcano)+
  geom_point(alpha=0.85,mapping= aes(
    fill=colors, x=lfc_CompostCOMP,y=-log(p_CompostCOMP)),
    size=5,color='black',pch=21) +
  scale_fill_manual(values = c('chocolate4','gray50'))+
  theme_classic()+
  labs(x=expression("log"["2"]~"Fold Change"))+
  theme(axis.text=element_text(size=10, colour = "black"), 
        axis.line = element_line(colour = "black"),
        axis.title.y=element_blank(),
        legend.position="none");p.fun.lab

ggsave(filename = "../Data/output/volcano.fun.lab.svg", plot = p.fun.lab,    width = 8, height = 8, device = "svg")

```

# Field trial

## Read in data
```{r, message=F, warning=F}
bacterial_otu_mat<- read_csv('../Data/Bloom_otu_table_Bac.csv',)
fungal_otu_mat<- read_csv('../Data/Bloom_otu_table_Fun.csv',)
bacterial_tax_mat<- read_csv('../Data/Bloom_taxonomy_table_Bac.csv',)
fungal_tax_mat<- read_csv('../Data/Bloom_taxonomy_table_Fun.csv',)
comb_samples_df <- read_csv('../Data/Bloom_sample_info.csv',)
```

## Create phyloseq and tse objects
```{r, message=F, warning=F}
bacterial_otu_mat <- bacterial_otu_mat %>%
  tibble::column_to_rownames("...1") 
fungal_otu_mat <- fungal_otu_mat %>%
  tibble::column_to_rownames("...1") 
bacterial_tax_mat <- bacterial_tax_mat %>%
  tibble::column_to_rownames("...1") 
fungal_tax_mat <- fungal_tax_mat %>%
  tibble::column_to_rownames("...1") 
comb_samples_df <- comb_samples_df %>%
  tibble::column_to_rownames("IMR.Label")

bacterial_tax_mat$Species <- if_else(is.na(bacterial_tax_mat$Species),NA,paste(bacterial_tax_mat$Genus,"_",bacterial_tax_mat$Species,sep = ""))
fungal_tax_mat$Species <- if_else(is.na(fungal_tax_mat$Species),NA,paste("s__",substr(fungal_tax_mat$Genus, 4, nchar(fungal_tax_mat$Genus)),"_",substr(fungal_tax_mat$Species, 4, nchar(fungal_tax_mat$Species)),sep = ""))

bacterial_tax_mat$Kingdom<-sub("^","k__",bacterial_tax_mat$Kingdom)
bacterial_tax_mat$Phylum<-sub("^","p__",bacterial_tax_mat$Phylum)
bacterial_tax_mat$Class<-sub("^","c__",bacterial_tax_mat$Class)
bacterial_tax_mat$Order<-sub("^","o__",bacterial_tax_mat$Order)
bacterial_tax_mat$Family<-sub("^","f__",bacterial_tax_mat$Family)
bacterial_tax_mat$Genus<-sub("^","g__",bacterial_tax_mat$Genus)
bacterial_tax_mat$Species<-sub("^","s__",bacterial_tax_mat$Species)

bacterial_tax_mat <- as.data.frame(bacterial_tax_mat)
fungal_tax_mat <- as.data.frame(fungal_tax_mat)

bacterial_tax_mat$Species[grep('incertae', bacterial_tax_mat$Species, ignore.case = TRUE)] <- NA
bacterial_tax_mat$Species[grep('unknown', bacterial_tax_mat$Species, ignore.case = TRUE)] <- NA
bacterial_tax_mat$Genus[grep('incertae', bacterial_tax_mat$Genus, ignore.case = TRUE)] <- NA
bacterial_tax_mat$Genus[grep('unknown', bacterial_tax_mat$Genus, ignore.case = TRUE)] <- NA
bacterial_tax_mat$Family[grep('incertae', bacterial_tax_mat$Family, ignore.case = TRUE)] <- NA
bacterial_tax_mat$Family[grep('unknown', bacterial_tax_mat$Family, ignore.case = TRUE)] <- NA
fungal_tax_mat$Species[grep('incertae', fungal_tax_mat$Species, ignore.case = TRUE)] <- NA
fungal_tax_mat$Species[grep('unknown', fungal_tax_mat$Species, ignore.case = TRUE)] <- NA
fungal_tax_mat$Genus[grep('incertae', fungal_tax_mat$Genus, ignore.case = TRUE)] <- NA
fungal_tax_mat$Genus[grep('unknown', fungal_tax_mat$Genus, ignore.case = TRUE)] <- NA
fungal_tax_mat$Family[grep('incertae', fungal_tax_mat$Family, ignore.case = TRUE)] <- NA
fungal_tax_mat$Family[grep('unknown', fungal_tax_mat$Family, ignore.case = TRUE)] <- NA

bac_combined_tax <- if_else(bacterial_tax_mat$Species == "_","test_true",bacterial_tax_mat$Species, missing = bacterial_tax_mat$Genus)
bac_combined_tax <- if_else(bac_combined_tax == "_","test_true",bac_combined_tax, missing = bacterial_tax_mat$Family)
bac_combined_tax <- if_else(bac_combined_tax == "_","test_true",bac_combined_tax, missing = bacterial_tax_mat$Order)
bac_combined_tax <- if_else(bac_combined_tax == "_","test_true",bac_combined_tax, missing = bacterial_tax_mat$Class)
bac_combined_tax <- if_else(bac_combined_tax == "_","test_true",bac_combined_tax, missing = bacterial_tax_mat$Phylum)
bac_combined_tax <- if_else(bac_combined_tax == "_","test_true",bac_combined_tax, missing = bacterial_tax_mat$Kingdom)
bac_combined_tax <- melt(data.frame(rownames(bacterial_tax_mat),bac_combined_tax))
colnames(bac_combined_tax) <- c("OTU", "taxon")

fun_combined_tax <- if_else(fungal_tax_mat$Species == "_","test_true",fungal_tax_mat$Species, missing = fungal_tax_mat$Genus)
fun_combined_tax <- if_else(fun_combined_tax == "_","test_true",fun_combined_tax, missing = fungal_tax_mat$Family)
fun_combined_tax <- if_else(fun_combined_tax == "_","test_true",fun_combined_tax, missing = fungal_tax_mat$Order)
fun_combined_tax <- if_else(fun_combined_tax == "_","test_true",fun_combined_tax, missing = fungal_tax_mat$Class)
fun_combined_tax <- if_else(fun_combined_tax == "_","test_true",fun_combined_tax, missing = fungal_tax_mat$Phylum)
fun_combined_tax <- if_else(fun_combined_tax == "_","test_true",fun_combined_tax, missing = fungal_tax_mat$Kingdom)
fun_combined_tax <- melt(data.frame(rownames(fungal_tax_mat),fun_combined_tax))
colnames(fun_combined_tax) <- c("OTU", "taxon")

comb_samples_df$Participant.ID <- as.factor(comb_samples_df$Participant.ID)
comb_samples_df$B.E_10_Compost <- as.factor(comb_samples_df$B.E_10_Compost)

comb_samples_df$IMR_label <- rownames(comb_samples_df)

bacterial_otu_mat <- as.matrix(bacterial_otu_mat)
fungal_otu_mat <- as.matrix(fungal_otu_mat)
bacterial_tax_mat <- as.matrix(bacterial_tax_mat)
fungal_tax_mat <- as.matrix(fungal_tax_mat)

OTU = otu_table(bacterial_otu_mat, taxa_are_rows = TRUE)
TAX = tax_table(bacterial_tax_mat)
samples = sample_data(comb_samples_df)

bacterial_phyloseq <- phyloseq(OTU, TAX, samples)
bacterial_phyloseq

OTU = otu_table(fungal_otu_mat, taxa_are_rows = TRUE)
TAX = tax_table(fungal_tax_mat)
samples = sample_data(comb_samples_df)

fungal_phyloseq <- phyloseq(OTU, TAX, samples)
fungal_phyloseq


###makes subset of bacterial and fungal samples for 20 paired composted/uncomposted sites
bac_compost_phyloseq <- subset_samples(bacterial_phyloseq, 
                                       Participant.ID =="132" |
                                         Participant.ID =="118"|
                                         Participant.ID =="156"|
                                         Participant.ID =="172"| 
                                         Participant.ID =="21"| 
                                         Participant.ID =="213"| 
                                         Participant.ID =="223"| 
                                         Participant.ID =="226"| 
                                         Participant.ID =="232"| 
                                         Participant.ID =="244")
sample_names(bac_compost_phyloseq)
fun_compost_phyloseq <- subset_samples(fungal_phyloseq, 
                                       Participant.ID =="132" | 
                                         Participant.ID =="118"| 
                                         Participant.ID =="156"| 
                                         Participant.ID =="172"| 
                                         Participant.ID =="21"| 
                                         Participant.ID =="213"| 
                                         Participant.ID =="223"| 
                                         Participant.ID =="226"| 
                                         Participant.ID =="232"| 
                                         Participant.ID =="244")
sample_names(fun_compost_phyloseq)
###makes tree summarized experiment objects from for each subset
bac_compost_tse = mia::convertFromPhyloseq(bac_compost_phyloseq)
fun_compost_tse = mia::convertFromPhyloseq(fun_compost_phyloseq)

```

## Run ancombc
```{r, message=F, warning=F}
###runs ancombc2 again for fungi and bacteria, but at the species level
#for bacteria
set.seed(123)
Sys.time()
compost_ancom_bac_SPE = ancombc2(data = bac_compost_tse, assay_name = "counts", tax_level = "Species",
                                 fix_formula = "B.E_10_Compost",
                                 p_adj_method = "holm", pseudo_sens = TRUE,
                                 prv_cut = 0.10, lib_cut = 0, s0_perc = 0.05,
                                 group = NULL, struc_zero = FALSE, neg_lb = FALSE,
                                 alpha = 0.05, n_cl = 4, verbose = TRUE,
                                 global = FALSE, pairwise = FALSE, dunnet = FALSE, trend = FALSE,
                                 iter_control = list(tol = 1e-2, max_iter = 20, verbose = TRUE),
                                 em_control = list(tol = 1e-5, max_iter = 100),
                                 lme_control = lme4::lmerControl(),
                                 mdfdr_control = list(fwer_ctrl_method = "holm", B = 100),
                                 trend_control = NULL)
Sys.time()
#for fungi
set.seed(123)
Sys.time() 
compost_ancom_fun_SPE = ancombc2(data = fun_compost_tse, assay_name = "counts", tax_level = "Species",
                                 fix_formula = "B.E_10_Compost",
                                 p_adj_method = "holm", pseudo_sens = TRUE,
                                 prv_cut = 0.10, lib_cut = 0, s0_perc = 0.05,
                                 group = NULL, struc_zero = FALSE, neg_lb = FALSE,
                                 alpha = 0.05, n_cl = 4, verbose = TRUE,
                                 global = FALSE, pairwise = FALSE, dunnet = FALSE, trend = FALSE,
                                 iter_control = list(tol = 1e-2, max_iter = 20, verbose = TRUE),
                                 em_control = list(tol = 1e-5, max_iter = 100),
                                 lme_control = lme4::lmerControl(),
                                 mdfdr_control = list(fwer_ctrl_method = "holm", B = 100),
                                 trend_control = NULL)
Sys.time()
```

### Save RDS
```{r, message=F, warning=F}
saveRDS(compost_ancom_bac_SPE, file = '../Data/output/Short_read_compost_ancom_bac_SPE.rds', ascii = FALSE, version = NULL,
        compress = TRUE, refhook = NULL)
saveRDS(compost_ancom_fun_SPE, file = '../Data/output/Short_read_compost_ancom_fun_SPE.rds', ascii = FALSE, version = NULL,
        compress = TRUE, refhook = NULL)


compost_ancom_bac_SPE <- readRDS(file = '../Data/output/Short_read_compost_ancom_bac_SPE.rds', refhook = NULL)
compost_ancom_fun_SPE <- readRDS(file = '../Data/output/Short_read_compost_ancom_fun_SPE.rds', refhook = NULL)
```

## Make figure for bacteria

### Get df ready for figure
```{r, message=F, warning=F}
res_prim = compost_ancom_bac_SPE$res

df_fig_compost_volcano = res_prim %>%
  dplyr::filter(p_B.E_10_Compost1 <1) %>%
  dplyr::arrange(desc(lfc_B.E_10_Compost1)) %>%
  dplyr::filter(passed_ss_B.E_10_Compost1 == 1) %>%
  dplyr::mutate(direct = ifelse(lfc_B.E_10_Compost1 > 0, "Positive LFC", "Negative LFC"))
df_fig_compost$taxon = factor(df_fig_compost$taxon, levels = df_fig_compost$taxon)
df_fig_compost$direct = factor(df_fig_compost$direct,
                               levels = c("Positive LFC", "Negative LFC"))


df_fig_compost_volcano$sig <- ifelse(df_fig_compost_volcano$p_B.E_10_Compost1>0.01, "NS",
                                       ifelse(df_fig_compost_volcano$p_B.E_10_Compost1==0.01, "NS",
                                              ifelse(df_fig_compost_volcano$p_B.E_10_Compost1<0.01, "sig",NA)))
df_fig_compost_volcano$sig.stage <- paste(df_fig_compost_volcano$direct,df_fig_compost_volcano$sig)


df_fig_compost_volcano$colors <- ifelse(df_fig_compost_volcano$sig=="NS", "NS",
                        ifelse(df_fig_compost_volcano$sig.stage=='Positive LFC sig',"Compost",
                               ifelse(df_fig_compost_volcano$sig.stage=='Negative LFC sig','No Compost',NA)))
df_fig_compost_volcano<- arrange(df_fig_compost_volcano, -log(p_B.E_10_Compost1))
```

### Make figure for bacteria
```{r, message=F, warning=F}
p.bac.field<-ggplot(data=df_fig_compost_volcano)+
  geom_point(alpha=0.85,mapping= aes(
    fill=colors, x=lfc_B.E_10_Compost1,y=-log(p_B.E_10_Compost1)),
    size=5,color='black',pch=21) +
  scale_fill_manual(values = c('chocolate4','gray50'))+
  theme_classic()+
  labs(y=expression("-log"["10"] ~ "p" ["adj"]), 
       x=expression("log"["2"]~"Fold Change"))+
  theme(axis.text=element_text(size=10, colour = "black"), 
        axis.line = element_line(colour = "black"),
        legend.position="none");p.bac.field
ggsave(filename = "../Data/output/volcano.bac.field.svg", plot = p.bac.field,
       width = 8, height = 8, device = "svg")


```

### Makes df for fungi
```{r, message=F, warning=F}
res_prim = compost_ancom_fun_SPE$res


df_fig_compost = res_prim %>%
  dplyr::filter(p_B.E_10_Compost1 <0.01) %>%
  dplyr::arrange(desc(lfc_B.E_10_Compost1)) %>%
  dplyr::filter(passed_ss_B.E_10_Compost1 == 1) %>%
  dplyr::mutate(direct = ifelse(lfc_B.E_10_Compost1 > 0, "Positive LFC", "Negative LFC"))
df_fig_compost$taxon = factor(df_fig_compost$taxon, levels = df_fig_compost$taxon)
df_fig_compost$direct = factor(df_fig_compost$direct, 
                               levels = c("Positive LFC", "Negative LFC"))


write.csv(df_fig_compost,"../Data/output/Short_read_fun_ancom_SPE_figure_data.csv")

df_fig_compost_volcano = res_prim %>%
  dplyr::filter(p_B.E_10_Compost1 <1) %>%
  dplyr::arrange(desc(lfc_B.E_10_Compost1)) %>%
  dplyr::filter(passed_ss_B.E_10_Compost1 == 1) %>%
  dplyr::mutate(direct = ifelse(lfc_B.E_10_Compost1 > 0, "Positive LFC", "Negative LFC"))
df_fig_compost$taxon = factor(df_fig_compost$taxon, levels = df_fig_compost$taxon)
df_fig_compost$direct = factor(df_fig_compost$direct,
                               levels = c("Positive LFC", "Negative LFC"))


df_fig_compost_volcano$sig <- ifelse(df_fig_compost_volcano$p_B.E_10_Compost1>0.01, "NS",
                                       ifelse(df_fig_compost_volcano$p_B.E_10_Compost1==0.01, "NS",
                                              ifelse(df_fig_compost_volcano$p_B.E_10_Compost1<0.01, "sig",NA)))
df_fig_compost_volcano$sig.stage <- paste(df_fig_compost_volcano$direct,df_fig_compost_volcano$sig)


df_fig_compost_volcano$colors <- ifelse(df_fig_compost_volcano$sig=="NS", "NS",
                        ifelse(df_fig_compost_volcano$sig.stage=='Positive LFC sig',"Compost",
                               ifelse(df_fig_compost_volcano$sig.stage=='Negative LFC sig','No Compost',NA)))
df_fig_compost_volcano<- arrange(df_fig_compost_volcano, -log(p_B.E_10_Compost1))
```


```{r, message=F, warning=F}
p.fun.field<-ggplot(data=df_fig_compost_volcano)+
  geom_point(alpha=0.85,mapping= aes(
    fill=colors, x=lfc_B.E_10_Compost1,y=-log(p_B.E_10_Compost1)),
    size=5,color='black',pch=21) +
  scale_fill_manual(values = c('chocolate4','cyan4','gray50'))+
  theme_classic()+
  labs(y=expression("-log"["10"] ~ "p" ["adj"]), 
       x=expression("log"["2"]~"Fold Change"))+
  theme(axis.text=element_text(size=10, colour = "black"), 
        axis.line = element_line(colour = "black"),
        legend.position="none");p.fun.field
ggsave(filename = "../Data/output/volcano.fun.field.svg", plot = p.fun.field,  width = 8, height = 8, device = "svg")
```

# Compile figure
```{r, message=F, warning=F}
grobs <- ggplotGrob(p.fun.field)$grobs
legend <- grobs[[which(sapply(grobs, function(x) x$name) == "guide-box")]]

fig6.nolegend <- plot_grid(p.bac.field, p.fun.field,p.bac.lab,p.fun.lab, ncol = 2, nrow = 2,labels = c("A", "B", "C","D"),label_size = 15, rel_heights = c(1,1,1,1));fig6.nolegend
fig6<-plot_grid(fig6.nolegend,legend, ncol=2, rel_widths=c(1,0.1));fig6

ggsave(filename = "../Data/output/volcano.fig6.svg", plot = fig6, width = 12, height = 8, device = "svg")
```

