---
title: "Fig_3_Stats"
author: "MD and EM"
date: "`r Sys.Date()`"
output: 
 html_document:
   toc: true
   toc_float: true
   editor_options:
   chunck_output_type: console
---
Inputs: <br>

-   EJM-CLS-22F_pct98.csv
-   EJM-CLS-22B_pct98.csv
-   Bacterial_alpha_div_data.csv
-   Fungal_alpha_div_data.csv
 

Outputs: <br>

-   figure 4


# Load libraries 
```{r, results = 'hide', message=FALSE}
knitr::opts_chunk$set(out.width = "100%")
library(cowplot)
library(tidyverse)
library(emmeans)
library(readxl)
library(phyloseq)
library(mia)
library(vegan)
library(pairwiseAdonis)
library(car)
library(agricolae)
```

# Read in data 
```{r}
FUN <- read.csv("../Data/EJM-CLS-22F_pct98.csv")
BAC <- read.csv("../Data/EJM-CLS-22B_pct98.csv")
BAC <- BAC[!is.na(BAC$Kingdom),]
FUN <- FUN[FUN$Kingdom =="k__Fungi",]
FUN <- FUN[!is.na(FUN$Kingdom),]
SAMPLE_DATA <- as.data.frame(read_xlsx("../Data/Sequencing_data_formatted_for_analysis(01.28.2026).xlsx", sheet = "Soil_sample_info"))

rownames(FUN) <- FUN$X
FUN <- FUN[,-1]
rownames(BAC) <- BAC$X
BAC <- BAC[,-1]

BAC$Kingdom<-sub("^","k__",BAC$Kingdom)
BAC$Kingdom<-sub("k__NA",NA,BAC$Kingdom)
BAC$Phylum<-sub("^","p__",BAC$Phylum)
BAC$Phylum<-sub("p__NA",NA,BAC$Phylum)
BAC$Class<-sub("^","c__",BAC$Class)
BAC$Class<-sub("c__NA",NA,BAC$Class)
BAC$Order<-sub("^","o__",BAC$Order)
BAC$Order<-sub("o__NA",NA,BAC$Order)
BAC$Family<-sub("^","f__",BAC$Family)
BAC$Family<-sub("f__NA",NA,BAC$Family)
BAC$Genus<-sub("^","g__",BAC$Genus)
BAC$Genus<-sub("g__NA",NA,BAC$Genus)
BAC$Species<-sub("^","s__",BAC$Species)
BAC$Species<-sub("s__NA",NA,BAC$Species)

SAMPLE_DATA$EJM_SOIL_NUMBER <- as.factor(SAMPLE_DATA$EJM_SOIL_NUMBER)

FUN_OTU<- FUN[,8:length(colnames(FUN))]
FUN_TAX <- FUN[,1:7]
BAC_OTU<- BAC[,8:length(colnames(BAC))]
BAC_TAX <- BAC[,1:7]

BAC_OTU <- BAC_OTU[rowSums(BAC_OTU)>1,]
FUN_OTU <- FUN_OTU[rowSums(FUN_OTU)>1,]

BAC_OTU <- BAC_OTU[,colSums(BAC_OTU)>1000]
FUN_OTU <- FUN_OTU[,colSums(FUN_OTU)>1000]

colSums(BAC_OTU)

rownames(SAMPLE_DATA)<- SAMPLE_DATA$SEQUENCED_INV_NUMBER
```

# Bacteria PCoA 

## Manage Data

### Generate phyloseq object

```{r}
SAMPLE_DATA_LAB <- SAMPLE_DATA[SAMPLE_DATA$SOIL_SOURCE=="LAB",]
SAMPLE_DATA_LAB <- SAMPLE_DATA_LAB[!is.na(SAMPLE_DATA_LAB$Collection_Date),]
SAMPLE_DATA_LAB <- SAMPLE_DATA_LAB[SAMPLE_DATA_LAB$Collection_Date!="45777",]
OTU <- BAC_OTU[,colnames(BAC_OTU) %in% rownames(SAMPLE_DATA_LAB)]
OTU <- OTU[rowSums(OTU)> 1,]

rowSums(OTU)

OTU <- as.matrix(OTU)
TAX = BAC_TAX[rownames(OTU),]
samples = SAMPLE_DATA_LAB[colnames(OTU),]
sample_names(samples) <- samples$SEQUENCED_INV_NUMBER
rownames(samples) <- samples$SEQUENCED_INV_NUMBER
em_tse <- TreeSummarizedExperiment(assays = list(Count = OTU),
                                   rowData = TAX,
                                   colData = samples)
assayNames(em_tse) <- "counts" 
```

### Model for permanova
```{r, message = F}
# Adonis - Mary code
P.bac.count <- convertToPhyloseq(em_tse,assay.type = "counts")
P.bac.rel = transform_sample_counts(P.bac.count, function(x) x / sum(x))
d <- data.frame(sample_data(P.bac.rel))
data <- data.frame(t(otu_table(P.bac.rel)))
data <- decostand(data, method='hellinger')

dist <- vegdist(data, method='bray')

adonis2(dist ~  Slurry*Compost, data=d, perm=999, by = "terms")
pairwise.adonis(dist,factors=d$Compost, p.adjust.m = "fdr")
pairwise.adonis(dist,factors=interaction(d$Compost,d$Slurry), p.adjust.m = "fdr")
```

## PCoA

### PCoA df
```{r, warning = F}
pcoa_results <- cmdscale(dist, k = 2, eig = TRUE, add = TRUE) 
pcoa_positions <- as_tibble(pcoa_results$points)
colnames(pcoa_positions) <- c("PCoA1", "PCoA2") 
points <- scores(pcoa_results)
points<- as.data.frame(points)
colnames(points) <- c("PCoA1", "PCoA2") 
points$SEQUENCED_INV_NUMBER <- rownames(points)

eigenvalues <- pcoa_results$eig
percent_explained_PC1 <- round(100 * eigenvalues[1] / sum(eigenvalues), 1)
percent_explained_PC2 <- round(100 * eigenvalues[2] / sum(eigenvalues), 1)
d <- data.frame(sample_data(P.bac.rel))
pcoa_data <- dplyr::inner_join(points,d, by = "SEQUENCED_INV_NUMBER")
pcoa_data$Year <- as.factor(pcoa_data$Year)
```

### PCoA Figure
```{r}
pcoa_data <- pcoa_data %>%
mutate(Compost = as_factor(Compost)) %>%
mutate(Compost = fct_relevel(Compost, "NONE", "COMP"))

p.bac.pcoa <- ggplot(data = pcoa_data, aes(x=PCoA1,y=PCoA2)) +
  geom_point(aes(color = interaction(Compost, Slurry)), size = 5) +
  labs(x = paste("PCoA 1 (", percent_explained_PC1, "%)", sep = ""),
    y = paste("PCoA 2 (", percent_explained_PC2, "%)", sep = "")) +
  stat_ellipse(aes(group= interaction(Compost, Slurry), color= interaction(Compost, Slurry)), type='t', level=0.95) +
  scale_color_discrete(palette = c("#EC7063","#B03A2E","#5DADE2","#21618C")) +
  theme_classic()+
  theme(legend.position = "none");p.bac.pcoa
```

# Fungi PCoA
## Manage Data

### Generate phyloseq object
```{r}
SAMPLE_DATA_LAB <- SAMPLE_DATA[SAMPLE_DATA$SOIL_SOURCE=="LAB",]
SAMPLE_DATA_LAB <- SAMPLE_DATA_LAB[!is.na(SAMPLE_DATA_LAB$Collection_Date),]
SAMPLE_DATA_LAB <- SAMPLE_DATA_LAB[SAMPLE_DATA_LAB$Collection_Date!="45777",]


OTU <- FUN_OTU[,colnames(FUN_OTU) %in% rownames(SAMPLE_DATA_LAB)]
OTU <- OTU[rowSums(OTU)> 1,]

OTU <- as.matrix(OTU)
TAX = FUN_TAX[rownames(OTU),]
samples = SAMPLE_DATA_LAB[colnames(OTU),]
sample_names(samples) <- samples$SEQUENCED_INV_NUMBER
rownames(samples) <- samples$SEQUENCED_INV_NUMBER
em_tse <- TreeSummarizedExperiment(assays = list(Count = OTU),
                                   rowData = TAX,
                                   colData = samples)
assayNames(em_tse) <- "counts" 

```

### Model for permanova
```{r, message = F}
P.fun.count <- convertToPhyloseq(em_tse,assay.type = "counts")
P.fun.rel = transform_sample_counts(P.fun.count, function(x) x / sum(x))
d <- data.frame(sample_data(P.fun.rel))
data <- data.frame(t(otu_table(P.fun.rel)))
data <- decostand(data, method='hellinger')

dist <- vegdist(data, method='bray')
adonis2(dist ~  Slurry*Compost, data=d, perm=999, by = "terms")
pairwise.adonis(dist,factors=d$Compost, p.adjust.m = "fdr")
pairwise.adonis(dist,factors=interaction(d$Compost,d$Slurry), p.adjust.m = "fdr")
```

## PCoA

## Make df for PCA

```{r, warning = F}
pcoa_results <- cmdscale(dist, k = 2, eig = TRUE, add = TRUE) 
pcoa_positions <- as_tibble(pcoa_results$points)
colnames(pcoa_positions) <- c("PCoA1", "PCoA2") 
points <- scores(pcoa_results)
points<- as.data.frame(points)
colnames(points) <- c("PCoA1", "PCoA2") 
points$SEQUENCED_INV_NUMBER <- rownames(points)


eigenvalues <- pcoa_results$eig
percent_explained_PC1 <- round(100 * eigenvalues[1] / sum(eigenvalues), 1)
percent_explained_PC2 <- round(100 * eigenvalues[2] / sum(eigenvalues), 1)
d <- data.frame(sample_data(P.fun.rel))
pcoa_data <- dplyr::inner_join(points,d, by = "SEQUENCED_INV_NUMBER")

pcoa_data <- pcoa_data %>%
mutate(Compost = as_factor(Compost)) %>%
mutate(Compost = fct_relevel(Compost, "NONE", "COMP"))
```

## PCoA Figure
```{r, message= F, warning=F}
p.fun.pcoa <- ggplot(data = pcoa_data, aes(x=PCoA1,y=PCoA2)) +
  geom_point(aes(color = interaction(Compost, Slurry)), size = 5) +
  labs(x = paste("PCoA 1 (", percent_explained_PC1, "%)", sep = ""),
    y = paste("PCoA 2 (", percent_explained_PC2, "%)", sep = "")) +
  stat_ellipse(aes(group= interaction(Compost, Slurry), color= interaction(Compost, Slurry)), type='t', level=0.95) +
  scale_color_discrete(palette = c("#EC7063","#B03A2E","#5DADE2","#21618C"), labels = c("Crossroads No Compost", "Crossroads With Compost", "Remembrance No Compost", "Remembrance With Compost")) +
  theme_classic()+
  theme(legend.position = "inside",
        legend.position.inside = c(0.8, 0.9),
        legend.title = element_blank());p.fun.pcoa
```

## Compiled PCoA
```{r, message= F, warning=F}
fig4.bottom <- plot_grid(p.bac.pcoa, p.fun.pcoa, labels = c('C', 'D'), rel_widths = c(1, 1));fig4.bottom
ggsave(filename = "../Data/output/fig4.pcoa.svg", plot = fig4.bottom, width = 11, height = 5, device = "svg")
```


# Alpha diversity

## Read in data 
```{r}
bac_alpha <- read.csv("../Data/Bacterial_alpha_div_data.csv")
fun_alpha <- read.csv("../Data/Fungal_alpha_div_data.csv")
```

## Run model for bacteria
```{r}
mod <- lm(shannon.Estimator~Slurry*Compost, data=bac_alpha)
par(mfrow=c(2,2))
plot(mod)
Anova(mod, Type = 3)
emmip(mod, Compost ~ Slurry, CIs = TRUE)
res <- emmeans(mod, list(pairwise ~ Compost*Slurry), adjust = "fdr")
plot(res[[1]], horizontal=F)

t <- HSD.test(mod, "Slurry");t
t <- HSD.test(mod, "Compost");t
```

## Make bacteria figure 

Generate df for figure
```{r, message=F}
bac_alpha$letters <- ifelse(bac_alpha$Slurry == "REM" & bac_alpha$Compost == "NONE", "b",  
                            ifelse(bac_alpha$Slurry == "REM" & bac_alpha$Compost == "COMP", "a",
                                   ifelse(bac_alpha$Slurry == "CROSS" & bac_alpha$Compost == "NONE", "b",
                                          ifelse(bac_alpha$Slurry == "CROSS" & bac_alpha$Compost == "COMP", "a",NA))))

bac_alpha$combined <- paste(bac_alpha$Slurry,bac_alpha$Compost)
bac_alpha$combined = factor(bac_alpha$combined, levels=c("CROSS NONE", "CROSS COMP", "REM NONE", "REM COMP"))
```

Make figure
```{r}
p.bac.alpha <- ggplot(bac_alpha, aes(combined, shannon.Estimator, fill = combined)) +
  geom_bar(stat = "summary", position = "dodge") +
  geom_errorbar(stat = "summary", position = "dodge", width=0.3, alpha=0.9, linewidth=0.8) +
  geom_jitter(width = 0.1, height = 0,color = "gray25", alpha = 0.6) +
  geom_text(stat = "summary", aes(label = letters),position=position_dodge(0.9), vjust = -2, hjust = -0.2, size = 5)+
  scale_fill_manual(values=c("#EC7063","#B03A2E","#5DADE2","#21618C")) +
  scale_x_discrete(labels=c('Crossroads\nNo Compost','Crossroads\nWith Compost','Remembrance\nNo Compost','Remembrance\nWith Compost'))+
  xlab(label = "") +
  ylab(label = "Asymptotic Shannon Diversity Estimate") +
  theme_classic()+
  theme(legend.position = "none");p.bac.alpha
```

## Run model for fungi
```{r}
mod <- lm(shannon.Estimator~Slurry*Compost, data=fun_alpha)
plot(mod)
Anova(mod, Type = 3)
emmip(mod, Compost ~ Slurry, CIs = TRUE)
res <- emmeans(mod, list(pairwise ~ Compost*Slurry), adjust = "fdr")
plot(res[[1]], horizontal=F)

t <- HSD.test(mod, "Slurry");t
t <- HSD.test(mod, "Compost");t
```

## Make figure for fungi

Generate df for figure
```{r, message= F, warning=F}
fun_alpha$letters <- ifelse(fun_alpha$Slurry == "REM" & fun_alpha$Compost == "NONE", "b",  
                            ifelse(fun_alpha$Slurry == "REM" & fun_alpha$Compost == "COMP", "a",
                                   ifelse(fun_alpha$Slurry == "CROSS" & fun_alpha$Compost == "NONE", "b",
                                          ifelse(fun_alpha$Slurry == "CROSS" & fun_alpha$Compost == "COMP", "a",NA))))

fun_alpha$combined <- paste(fun_alpha$Slurry,fun_alpha$Compost)
fun_alpha$combined = factor(fun_alpha$combined, levels=c("CROSS NONE", "CROSS COMP", "REM NONE", "REM COMP"))
```

Generate figure for fungi
```{r, message= F, warning=F}
p.fun.alpha <- ggplot(fun_alpha, aes(combined, shannon.Estimator, fill = combined)) +
  geom_bar(stat = "summary", position = "dodge") +
  geom_errorbar(stat = "summary", position = "dodge", width=0.3, alpha=0.9, linewidth=0.8) +
  geom_jitter(width = 0.1, height = 0,color = "gray25", alpha = 0.6) +
  geom_text(stat = "summary", aes(label = letters),position=position_dodge(0.9), vjust = -3, hjust = -0.2, size = 5)+
  scale_fill_manual(values=c("#EC7063","#B03A2E","#5DADE2","#21618C")) +
  scale_x_discrete(labels=c('Crossroads\nNo Compost','Crossroads\nWith Compost','Remembrance\nNo Compost','Remembrance\nWith Compost'))+
  xlab(label = "") +
  theme_classic()+
  theme(legend.position = "none",
        axis.title.y = element_blank());p.fun.alpha
```

# Compile all figs 
```{r, message= F, warning=F}
fig4.top<- plot_grid(p.bac.alpha,p.fun.alpha, labels = c("A","B")); fig4.top

fig4 <- plot_grid(fig4.top, fig4.bottom, ncol = 1,nrow = 2);fig4

ggsave(filename = "../Data/output/fig4.svg", plot = fig4,  width = 10, height = 8, device = "svg")
```













```



